Towards Practical First-Order Model Counting

### 2024-04-16 doc/paper.tex
Writing and stuff (+ some figures to describe smoothing)

\DeclareMathOperator{\Dom}{Dom}
\DeclareMathOperator{\Doms}{Doms}
\DeclareMathOperator{\Vars}{Vars}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator{\GDR}{GDR}

@@

\author{%
Ananth K. Kidambi$^1$\and
Guramrit Singh$^1$\and
Paulius Dilkas$^2$\and
Kuldeep S. Meel$^3$ \\
\affiliations
$^1$Indian Institute of Technology Bombay, Mumbai, India\\
$^2$National University of Singapore, Singapore, Singapore\\
$^3$University of Toronto, Toronto, Canada\\
\emails
\{210051002, 210050061\}@iitb.ac.in,
paulius.dilkas@nus.edu.sg,
meel@cs.toronto.edu
}

@@

\begin{abstract}
\end{abstract}

@@

\paragraph{Papers to cite.}
\begin{itemize}
  \item overviews
        \begin{itemize}
          \item lifted probabilistic inference~\cite{DBLP:conf/ecai/Kersting12}
          \item recent overview paper~\cite{DBLP:conf/ijcai/Kuzelka23}
        \end{itemize}
  \item Alternative definition~\cite{DBLP:journals/cacm/GogateD16}
  \item relevant theoretical work~\cite{DBLP:conf/aaai/MalhotraS22}
  \item original domain recursion~\cite{DBLP:conf/nips/Broeck11}
  \item algorithms
        \begin{itemize}
          \item \textsc{ForcLift}~\cite{DBLP:conf/ijcai/BroeckTMDR11}
          \item \textsc{Crane}~\cite{DBLP:conf/kr/DilkasB23}
          \item \textsc{FastWFOMC}~\cite{DBLP:conf/uai/BremenK21}
          \item \textsc{L2C}~\cite{DBLP:conf/kr/KazemiP16} (similarly to us
                compiles to C++ code, but (probably) doesn't work on as many
                formulas)
          \item approximate~\cite{DBLP:conf/ijcai/BremenK20}
          \item for Markov logic networks~\cite{DBLP:journals/ml/RichardsonD06}
                \begin{itemize}
                  \item \textsc{Magician}~\cite{DBLP:conf/aaai/VenugopalSG15}
                  \item \textsc{Tuffy}~\cite{DBLP:journals/pvldb/NiuRDS11}
                  \item \textsc{Alchemy}~\cite{DBLP:journals/cacm/GogateD16}
                        (same as the alternative definition)
                \end{itemize}
        \end{itemize}
  \item complexity
        \begin{itemize}
          \item liftability~\cite{DBLP:conf/starai/JaegerB12}
          \item hardness for three variables~\cite{DBLP:conf/pods/BeameBGS15}
          \item liftable fragments
                \begin{itemize}
                  \item \Ctwo~\cite{DBLP:journals/jair/Kuzelka21}
                  \item tree axioms~\cite{DBLP:journals/ai/BremenK23}
                  \item linear order axioms~\cite{DBLP:conf/aaai/TothK23}
                  \item some liftable fragments~\cite{DBLP:conf/nips/KazemiKBP16}
                \end{itemize}
        \end{itemize}
  \item applications
        \begin{itemize}
          \item extensions to
                sampling~\cite{DBLP:conf/aaai/WangB0K22,DBLP:conf/lics/WangP0K23}
          \item discovery of combinatorial
                sequences~\cite{DBLP:conf/ijcai/SvatosJT0K23}
          \item conjecturing recurrence
                relations~\cite{DBLP:conf/ilp/BarvinekB0ZK21}
          \item probabilistic logic
                programming~\cite{DBLP:journals/ijar/RiguzziBZCL17} (WFOMC was
                shown to be supreme)
          \item probabilistic databases~\cite{DBLP:journals/debu/GribkoffSB14}
        \end{itemize}
  \item lifted inference elsewhere
        \begin{itemize}
          \item constraint satisfaction~\cite{DBLP:journals/jair/TotisDRK23}
          \item answer set programming~\cite{DBLP:journals/ijar/AzzoliniR23}
        \end{itemize}
\end{itemize}

@@

\paragraph{Contributions (outdated).}
\begin{itemize}
  \item Completing the definitions of recursive functions by:
  \begin{itemize}
    \item identifying a sufficient set of base cases (\cref{sec:identifying})
    \item constructing formulas that correspond to these base cases
          (\cref{sec:simplifying})
    \item and recursing on these subproblems
  \end{itemize}
  \item Compiling these function definitions into a C++ program that can be
        executed independently for any domain size values \cref{sec:cpp}
        \begin{itemize}
          \item including support for infinite precision arithmetic via GNU
                Multiple Precision Arithmetic Library
        \end{itemize}
  \item Experiments comparing \Cranetwo{} with the main alternative approach
        demonstrate the ability of \Cranetwo{} to scale to domain sizes\dots
\end{itemize}

@@

\subsection{Logic}

@@

\paragraph{Three types of logics}
\begin{itemize}
  \item See \cref{tbl:logics} for a detailed comparison. The notation introduced
        in the table is standard for \Ctwo{}, new for \UFO{}, and redefined to
        be more specific for \FO{}.
  \item All three logics are function-free.
  \item Domains are always assumed to be finite.
  \item In many-sorted logic, each term is assigned to a \emph{sort}, and each
        predicate $p/n$ is assigned to a sequence of $n$ sorts. Each sort has
        its corresponding domain. In the input formula, all domains are assumed
        to be pairwise disjoin.t Most of these assignments are typically left
        implicit and can be reconstructed from the quantifiers. For instance,
        $\forall x,y \in \Delta\text{. }P(x, y)$ implies that variables $x$ and
        $y$ have the same sort. On the other hand, $\forall x \in \Delta\text{.
        }\forall y \in \Gamma\text{. } P(x, y)$ implies that $x$ and $y$ have
        different sorts, and it would be improper to have $x = y$ as part of a
        formula.
  \item
        \begin{itemize}
          \item \FO{} is used as the input format for
                \textsc{ForcLift}\footnote{\url{https://github.com/UCLA-StarAI/Forclift}}~\cite{DBLP:conf/ijcai/BroeckTMDR11}
                and its extensions
                \textsc{Crane}\footnote{\url{https://doi.org/10.5281/zenodo.8004077}}~\cite{DBLP:conf/kr/DilkasB23}
                and \Cranetwo{}.
          \item \Ctwo{} is discussed in the literature on
                \textsc{FastWFOMC}~\cite{DBLP:conf/uai/BremenK21} and related
                methods~\cite{DBLP:journals/jair/Kuzelka21,DBLP:conf/aaai/MalhotraS22}
          \item \UFO{} is the input format supported by a version of
                \textsc{FastWFOMC} obtained directly from the authors. Note that
                the publicly available
                version\footnote{\url{https://comp.nus.edu.sg/~tvanbr/software/fastwfomc.tar.gz}}
                does not support any cardinality constraints.
        \end{itemize}
  \item Note that, in the case of \textsc{ForcLift} and its extensions, support
        for a formula as valid input does not imply that the algorithm will be
        able to compile the formula into a circuit or graph suitable for lifted
        model counting. However, it is known that \textsc{ForcLift} compilation
        is guaranteed to succeed on any \FO{} formula without constants and with
        at most two variables~\cite{DBLP:conf/nips/Broeck11}.
\end{itemize}

\todo[inline,caption={}]{
  \begin{itemize}
    \item Add examples of cardinality constraints and counting quantifiers
    \item Mention the role of Skolemization (with a citation)
  \end{itemize}
}

\paragraph{The \FO{} logic (after Skolemization).}
\begin{itemize}
  \item In the spirit of keeping sorts implicit, we always assume formulas `type
        check' with respect to sorts. For example, if $P(x)$, $P(y)$, and
        $x \ne c$ are all part of the formula (for some predicate $P$, variables
        $x$ and $y$, and constant $c$), then $x$, $y$, and $c$ all have the same
        sort.
  \item A \emph{formula} is a conjunction of clauses.
  \item A \emph{clause} is of the form $\forall x_{1} \in \Delta_{1}\text{.
        }\forall x_{2} \in \Delta_{2}\dots\text{
        }\forall x_{n} \in \Delta_{n}\text{. }\phi(x_{1}, x_{2}, \dots, x_{n})$,
        where $\phi$ is a disjunction of literals that only contain variables
        $x_{1}, \dots, x_{n}$ (and any constants).
  \item A \emph{literal} is either an atom or its negation.
  \item An \emph{atom} is either:
  \begin{itemize}
    \item $P(t_{1}, \dots, t_{m})$ for some predicate $P/m$ and terms
          $t_{1}, \dots, t_{m}$ or
    \item $x=y$ for some terms $x$ and $y$
  \end{itemize}
  \item An atom is \emph{ground} if it contains no variables (i.e., only
        constants).
  \item The \emph{arity} of a predicate is the number of arguments it takes,
        i.e., $m$ in the case of predicate $P$.
  \item When we want to denote a predicate together with its arity, we write
        $P/m$.
  \item A \emph{term} is either a variable or a constant.
  \item Throughout the paper, we use set-theoretic notation, interpreting a
        formula as a set of clauses and a clause as a set of literals. Moreover,
        for readability, clauses written on separate lines are implicitly
        conjoined.
\end{itemize}

@@

\subsection{Algebra}

\todo[inline,caption={}]{
  \begin{itemize}
    \item check if I'm still using this notation in all these cases
    \item explain how this works for function calls too
    \item introduce the Iverson bracket (somewhere)
    \item reorder the Logic subsection to cover \FO before comparing it to other
          logics (and clarify the differences between the input format and the
          internal format)
  \end{itemize}
}

@@

\begin{definition}
  An \emph{equation} is always of the form $f(\mathbf{x}) = \expr{}$, where
  $f(\mathbf{x})$ is a signature, and $\expr{}$ is an algebraic expression.
  Henceforth, we call $f(\mathbf{x})$ and $\expr{}$ the left-hand side (LHS) and
  the right-hand side (RHS) of the equation, respectively.
\end{definition}

@@

\paragraph{Notation.}
We write $\expr{}$ for an arbitrary algebraic expression. In the context of both
algebra and logic, we write $C[x \mapsto y]$ for $C$ with all occurrences of $x$
replaced with $y$.

@@

@@

@@

\todo[inline]{Insert motivation for smoothing from Section 3.4.\ of the ForcLift
  paper.}

Originally, smoothing was (and still is) a two-step process. First, atoms that
are still accounted for in the circuit are propagated upwards. Then, at vertices
of certain types, missing atoms are detected and additional sinks are created to
account for them. If left unchanged, the first step of this process would result
in an infinite loop whenever a cycle is encountered. \Cref{alg:smoothing}
outlines how the first step can be adapted to an arbitrary directed graph.

@@

\paragraph{Benchmarks (probably for supplementary material).}
\begin{itemize}
  \item Functions
        \begin{itemize}
          \item In \Ctwo: $\forall x \in \Delta\text{.
                }\exists^{=1} y \in \Delta\text{. }P(x, y)$
          \item In \UFO:
                \begin{gather*}
                  \forall x, y \in \Delta\text{. }S(x) \lor \neg P(x, y)\\
                  |P| = |\Delta|
                \end{gather*}
          \item In \FO:
                \begin{gather*}
                  \forall x \in \Delta\text{. }\exists y \in \Delta\text{. }P(x, y)\\
                  \forall x, y, z \in \Delta\text{. }P(x, y) \land P(x, z) \Rightarrow y = z
                \end{gather*}
        \end{itemize}
  \item Permutations
        \begin{itemize}
          \item In \Ctwo:
                \begin{gather*}
                  \forall x \in \Delta\text{. }\exists^{=1} y \in \Delta\text{. }P(x, y)\\
                  \forall y \in \Delta\text{. }\exists^{=1} x \in \Delta\text{. }P(x, y)
                \end{gather*}
          \item In \UFO:
                \begin{gather*}
                  \forall x, y \in \Delta\text{. }R(x) \lor \neg P(x, y)\\
                  \forall x, y \in \Delta\text{. }S(x) \lor \neg P(y, x)\\
                  |P| = |\Delta|
                \end{gather*}
                with weights $w^{-}(R) = w^{-}(S) = -1$
          \item In \FO:
                \begin{gather*}
                  \forall x \in \Delta\text{. }\exists y \in \Delta\text{. }P(x, y)\\
                  \forall y \in \Delta\text{. }\exists x \in \Delta\text{. }P(x, y)\\
                  \forall x, y, z \in \Delta\text{. }P(x, y) \land P(x, z) \Rightarrow y = z\\
                  \forall x, y, z \in \Delta\text{. }P(x, y) \land P(z, y) \Rightarrow x = z
                \end{gather*}
        \end{itemize}
\end{itemize}

@@

@@

% TODO: the report has some (long) examples of formulas being transformed into
% programs perhaps suitable for supplementary material

>>>>

\DeclareMathOperator{\CR}{CR}
\DeclareMathOperator{\DR}{DR}
\DeclareMathOperator{\Reff}{Ref}

\DeclareMathOperator{\Dom}{Dom}
\DeclareMathOperator{\Doms}{Doms}
\DeclareMathOperator{\Vars}{Vars}
\DeclareMathOperator*{\argmin}{argmin}

@@

\author{%
  Ananth K. Kidambi$^1$\footnote{The first and second authors contributed equally and were affiliated with the National University of Singapore during the completion of this work.}\and
Guramrit Singh$^1$\and
Paulius Dilkas$^2$\and
Kuldeep S. Meel$^3$ \\
\affiliations
$^1$Indian Institute of Technology Bombay, Mumbai, India\\
$^2$National University of Singapore, Singapore, Singapore\\
$^3$University of Toronto, Toronto, Canada\\
\emails
\{210051002, 210050061\}@iitb.ac.in,
paulius.dilkas@nus.edu.sg,
meel@cs.toronto.edu
}

@@

\begin{abstract}
  TODO: at most 200 words
\end{abstract}

@@

\paragraph{Papers To Cite}
\begin{itemize}
  \item overviews
        \begin{itemize}
          \item lifted probabilistic inference~\cite{DBLP:conf/ecai/Kersting12}
          \item recent overview paper~\cite{DBLP:conf/ijcai/Kuzelka23}
        \end{itemize}
  \item Alternative definition~\cite{DBLP:journals/cacm/GogateD16}
  \item relevant theoretical work~\cite{DBLP:conf/aaai/MalhotraS22}
  \item original domain recursion~\cite{DBLP:conf/nips/Broeck11}
  \item algorithms
        \begin{itemize}
          \item \textsc{ForcLift}~\cite{DBLP:conf/ijcai/BroeckTMDR11}
          \item \textsc{Crane}~\cite{DBLP:conf/kr/DilkasB23}
          \item \textsc{FastWFOMC}~\cite{DBLP:conf/uai/BremenK21}
          \item \textsc{L2C}~\cite{DBLP:conf/kr/KazemiP16} (similarly to us
                compiles to C++ code, but (probably) doesn't work on as many
                formulas)
          \item approximate~\cite{DBLP:conf/ijcai/BremenK20}
          \item for Markov logic networks~\cite{DBLP:journals/ml/RichardsonD06}
                \begin{itemize}
                  \item \textsc{Magician}~\cite{DBLP:conf/aaai/VenugopalSG15}
                  \item \textsc{Tuffy}~\cite{DBLP:journals/pvldb/NiuRDS11}
                  \item \textsc{Alchemy}~\cite{DBLP:journals/cacm/GogateD16}
                        (same as the alternative definition)
                \end{itemize}
        \end{itemize}
  \item complexity
        \begin{itemize}
          \item liftability~\cite{DBLP:conf/starai/JaegerB12}
          \item hardness for three variables~\cite{DBLP:conf/pods/BeameBGS15}
          \item liftable fragments
                \begin{itemize}
                  \item \Ctwo~\cite{DBLP:journals/jair/Kuzelka21}
                  \item tree axioms~\cite{DBLP:journals/ai/BremenK23}
                  \item linear order axioms~\cite{DBLP:conf/aaai/TothK23}
                  \item some liftable fragments~\cite{DBLP:conf/nips/KazemiKBP16}
                \end{itemize}
        \end{itemize}
  \item applications
        \begin{itemize}
          \item extensions to
                sampling~\cite{DBLP:conf/aaai/WangB0K22,DBLP:conf/lics/WangP0K23}
          \item discovery of combinatorial
                sequences~\cite{DBLP:conf/ijcai/SvatosJT0K23}
          \item conjecturing recurrence
                relations~\cite{DBLP:conf/ilp/BarvinekB0ZK21}
          \item probabilistic logic
                programming~\cite{DBLP:journals/ijar/RiguzziBZCL17} (WFOMC was
                shown to be supreme)
          \item probabilistic databases~\cite{DBLP:journals/debu/GribkoffSB14}
        \end{itemize}
  \item lifted inference elsewhere
        \begin{itemize}
          \item constraint satisfaction~\cite{DBLP:journals/jair/TotisDRK23}
          \item answer set programming~\cite{DBLP:journals/ijar/AzzoliniR23}
        \end{itemize}
\end{itemize}

@@

\paragraph{Contributions}
\begin{itemize}
  \item Completing the definitions of recursive functions by:
  \begin{itemize}
    \item identifying a sufficient set of base cases (\cref{sec:identifying})
    \item constructing formulas that correspond to these base cases
          (\cref{sec:simplifying})
    \item and recursing on these subproblems
  \end{itemize}
  \item Compiling these function definitions into a C++ program that can be
        executed independently for any domain size values \cref{sec:cpp}
        \begin{itemize}
          \item including support for infinite precision arithmetic via GNU
                Multiple Precision Arithmetic Library
        \end{itemize}
  \item Experiments comparing \Cranetwo{} with the main alternative approach
        demonstrate the ability of \Cranetwo{} to scale to domain sizes\dots
\end{itemize}

@@

\Cref{sec:logic} introduces the fundamental ideas of first-order logic and
WFOMC, and discusses the different logics used in the latter.\@
\Cref{sec:algebra} introduces the terminology we use to describe the output of
the original \textsc{Crane} algorithm~\cite{DBLP:conf/kr/DilkasB23}, i.e.,
functions and equations that define them.

\subsection{Logic}\label{sec:logic}

\begin{example}\label{example:functions}
  Functions
  \begin{itemize}
    \item In \Ctwo: $\forall x \in \Delta\text{.
          }\exists^{=1} y \in \Delta\text{. }P(x, y)$
    \item In \UFO:
          \begin{gather*}
            \forall x, y \in \Delta\text{. }S(x) \lor \neg P(x, y)\\
            |P| = |\Delta|
          \end{gather*}
    \item In \FO:
          \begin{gather*}
            \forall x \in \Delta\text{. }\exists y \in \Delta\text{. }P(x, y)\\
            \forall x, y, z \in \Delta\text{. }P(x, y) \land P(x, z) \Rightarrow y = z
          \end{gather*}
  \end{itemize}
\end{example}

@@

\paragraph{Three Types of Logics}
\begin{itemize}
  \item See \cref{tbl:logics} for a detailed comparison. The notation introduced
        in the table is standard for \Ctwo{}, new for \UFO{}, and redefined to
        be more specific for \FO{}.
  \item All three logics are function-free.
  \item Domains are always assumed to be finite.
  \item In many-sorted logic, each term is assigned to a \emph{sort}, and each
        predicate $p/n$ is assigned to a sequence of $n$ sorts. Each sort has
        its corresponding domain. In the input formula, all domains are assumed
        to be pairwise disjoin.t Most of these assignments are typically left
        implicit and can be reconstructed from the quantifiers. For instance,
        $\forall x,y \in \Delta\text{. }P(x, y)$ implies that variables $x$ and
        $y$ have the same sort. On the other hand, $\forall x \in \Delta\text{.
        }\forall y \in \Gamma\text{. } P(x, y)$ implies that $x$ and $y$ have
        different sorts, and it would be improper to have $x = y$ as part of a
        formula.
  \item
        \begin{itemize}
          \item \FO{} is used as the input format for
                \textsc{ForcLift}\footnote{\url{https://github.com/UCLA-StarAI/Forclift}}~\cite{DBLP:conf/ijcai/BroeckTMDR11}
                and its extensions
                \textsc{Crane}\footnote{\url{https://doi.org/10.5281/zenodo.8004077}}~\cite{DBLP:conf/kr/DilkasB23}
                and \Cranetwo{}.
          \item \Ctwo{} is discussed in the literature on
                \textsc{FastWFOMC}~\cite{DBLP:conf/uai/BremenK21} and related
                methods~\cite{DBLP:journals/jair/Kuzelka21,DBLP:conf/aaai/MalhotraS22}
          \item \UFO{} is the input format supported by a version of
                \textsc{FastWFOMC} obtained directly from the authors. Note that
                the publicly available
                version\footnote{\url{https://comp.nus.edu.sg/~tvanbr/software/fastwfomc.tar.gz}}
                does not support any cardinality constraints.
        \end{itemize}
  \item Note that, in the case of \textsc{ForcLift} and its extensions, support
        for a formula as valid input does not imply that the algorithm will be
        able to compile the formula into a circuit or graph suitable for lifted
        model counting. However, it is known that \textsc{ForcLift} compilation
        is guaranteed to succeed on any \FO{} formula without constants and with
        at most two variables~\cite{DBLP:conf/nips/Broeck11}.
\end{itemize}

\paragraph{The \FO{} Logic (After Skolemization)}
\begin{itemize}
  \item Although existential quantifiers are supported, here we describe the
        format used internally. During preprocessing, all existential
        quantifiers are eliminated using
        Skolemization~\cite{DBLP:conf/kr/BroeckMD14}, and the formula is
        rewritten into a conjunction of clauses, each of which is in
        \emph{prenex normal form}~\cite{hinman2018fundamentals}.
  \item In the spirit of keeping sorts implicit, we always assume formulas `type
        check' with respect to sorts. For example, if $P(x)$, $P(y)$, and
        $x \ne c$ are all part of the formula (for some predicate $P$, variables
        $x$ and $y$, and constant $c$), then $x$, $y$, and $c$ all have the same
        sort.
  \item A \emph{formula} is a conjunction of clauses.
  \item A \emph{clause} is of the form $\forall x_{1} \in \Delta_{1}\text{.
        }\forall x_{2} \in \Delta_{2}\dots\text{
        }\forall x_{n} \in \Delta_{n}\text{. }\phi(x_{1}, x_{2}, \dots, x_{n})$,
        where $\phi$ is a disjunction of literals that only contain variables
        $x_{1}, \dots, x_{n}$ (and any constants).
  \item A \emph{literal} is either an atom or its negation.
  \item An \emph{atom} is either:
  \begin{itemize}
    \item $P(t_{1}, \dots, t_{m})$ for some predicate $P/m$ and terms
          $t_{1}, \dots, t_{m}$ or
    \item $x=y$ for some terms $x$ and $y$
  \end{itemize}
  \item An atom is \emph{ground} if it contains no variables (i.e., only
        constants).
  \item The \emph{arity} of a predicate is the number of arguments it takes,
        i.e., $m$ in the case of predicate $P$.
  \item When we want to denote a predicate together with its arity, we write
        $P/m$.
  \item A \emph{term} is either a variable or a constant.
  \item Throughout the paper, we use set-theoretic notation, interpreting a
        formula as a set of clauses and a clause as a set of literals. Moreover,
        for readability, clauses written on separate lines are implicitly
        conjoined.
\end{itemize}

@@

\subsection{Algebra}\label{sec:algebra}

\todo[inline,caption={}]{
  \begin{itemize}
    \item explain how this works for function calls too
    \item introduce the Iverson bracket (somewhere)
    \item reorder the Logic subsection to cover \FO{} before comparing it to
          other logics (and clarify the differences between the input format and
          the internal format)
    \item (somewhere) mention the trick about increasing the number of sorts
  \end{itemize}
}

@@

\paragraph{Additional notation}
We write $\expr{}$ for an arbitrary algebraic expression. In the context of both
algebra and logic, we write $C[x \mapsto y]$ for $C$ with all occurrences of $x$
replaced with $y$.

\begin{definition}
  An \emph{equation} is $f(\mathbf{x}) = \expr{}$, where $f(\mathbf{x})$ is a
  signature, and $\expr{}$ is an algebraic expression. Henceforth, we call
  $f(\mathbf{x})$ and $\expr{}$ the left-hand side (LHS) and the right-hand side
  (RHS) of the equation, respectively.
\end{definition}

@@

@@

\todo[inline]{Can I use an example based on permutations or something similar?}

@@

\todo[inline,caption={}]{
  \begin{itemize}
    \item It seems like I'll need to introduce:
          \begin{itemize}
            \item FCGs
            \item node types (be consistent w.r.t.\ node vs vertex)
            \item their notation
          \end{itemize}
    \item Describe \cref{fig:smoothing1,fig:smoothing2}
  \end{itemize}
}

\begin{figure}
  \centering
  \begin{forest}
    for tree={sn edges}
    [$\DR(c \in \Delta)$,name=dr,ellipse,draw,label={right:\textcolor{blue}{3. $\{\, \forall x \in \Delta\text{. }P(x) \,\}$}}
    [$\land$,ellipse,draw,label={left:\textcolor{blue}{2. $\{\, P(c) \,\}$}},label={right:\textcolor{blue}{6. $\{\, P(c), \forall x \in \Delta\text{. }x \ne c \Rightarrow P(x) \,\}$}},
    [$P(c)$,rectangle,draw,fill=green!20,label={272:\textcolor{blue}{1. $\{\, P(c) \,\}$}}],
    [$\CR(\Delta' \gets \Delta \setminus \{\, c \,\})$,ellipse,draw,label={87:\textcolor{blue}{5. $\{\, \forall x \in \Delta\text{. }x \ne c \Rightarrow P(x) \,\}$}}
    [$\Reff(\Delta \mapsto \Delta')$,name=ref,ellipse,draw,label={below:\textcolor{blue}{4. $\{\, \forall x \in \Delta'\text{. }P(x) \,\}$}}]
    ]
    ]
    ]
  \end{forest}
  \caption{An example based on permutation counting (but simplified a lot). The
    arc from the $\Reff$ node to the $\DR$ node is omitted.}\label{fig:smoothing1}
\end{figure}

\begin{figure}
  \centering
  \begin{forest}
    for tree={sn edges}
    [$\DR(c \in \Delta)$,name=dr,ellipse,draw,
    [$P(c)$,rectangle,draw,fill=green!20],
    [$\forall x \in \Delta\text{. }x \ne c \Rightarrow P(x) \lor \neg P(x)$,rectangle,dashed,edge=dashed,draw,fill=blue!20]
    ]
  \end{forest}
  \caption{An artificial example of a situation when smoothing nodes need to be
    added below a $\DR$ node}\label{fig:smoothing2}
\end{figure}

@@

Smoothing is a two-step process. First, atoms that are still accounted for in
the circuit are propagated upwards. Then, at vertices of certain types, missing
atoms are detected and additional sinks are created to account for them. If left
unchanged, the first step of this process would result in an infinite loop
whenever a cycle is encountered. \Cref{alg:smoothing} outlines how the first
step can be adapted to an arbitrary directed graph.

@@

\paragraph{Benchmarks (probably for supplementary material).}
\begin{itemize}
  \item Functions (\cref{example:functions})
  \item Permutations
        \begin{itemize}
          \item In \Ctwo:
                \begin{gather*}
                  \forall x \in \Delta\text{. }\exists^{=1} y \in \Delta\text{. }P(x, y)\\
                  \forall y \in \Delta\text{. }\exists^{=1} x \in \Delta\text{. }P(x, y)
                \end{gather*}
          \item In \UFO:
                \begin{gather*}
                  \forall x, y \in \Delta\text{. }R(x) \lor \neg P(x, y)\\
                  \forall x, y \in \Delta\text{. }S(x) \lor \neg P(y, x)\\
                  |P| = |\Delta|
                \end{gather*}
                with weights $w^{-}(R) = w^{-}(S) = -1$
          \item In \FO:
                \begin{gather*}
                  \forall x \in \Delta\text{. }\exists y \in \Delta\text{. }P(x, y)\\
                  \forall y \in \Delta\text{. }\exists x \in \Delta\text{. }P(x, y)\\
                  \forall x, y, z \in \Delta\text{. }P(x, y) \land P(x, z) \Rightarrow y = z\\
                  \forall x, y, z \in \Delta\text{. }P(x, y) \land P(z, y) \Rightarrow x = z
                \end{gather*}
          \item NOTE: when it's the same domain, counting bijections,
                injections, surjections, and partial surjections all boil down
                to the same integer sequence.
        \end{itemize}
\end{itemize}

\todo[inline]{ If it is possible to rewrite a formula to have more sorts, we do
  that. Explain in more detail and with an example. Mention that this has been
  observed previously by my previous work (i.e., linear vs cubic complexity). }

@@

\todo[inline,caption={}]{
  \begin{itemize}
    \item Maybe add some complete examples of C++ programs in the supplementary
          material.
          \item Later on:
          \begin{itemize}
            \item Eliminate (e.g., Emacs) warnings.
            \item must run it by A\&G
            \item re-check submission instructions and formatting guidelines
            \item Full stop at the end of a single-sentence caption?
          \end{itemize}
  \end{itemize}
}

\clearpage % TODO: temporarary

@@
